Distributed Topology Discovery in Self-Assembled Nano Network-On-Chip

### 2014-09-10 trunk/PAPER/introduction.tex
Fixed bib

Exploring long term alternatives to the CMOS technology is gaining
more and more relevance as the scaling trend of such devices keeps
introducing new challenges: power density, defect tolerance, testing
costs and wire delays are only a few of the many critical aspects
involved~\cite{itrs13}. 
While software parallelism and multicore
approaches~\cite{horowitz2004,powell2009} are partially mitigating the impact of such
constraints on performances, it is likely that a growing computing demand will
eventually need even more radical architectural modifications and 
paradigms to address the Computer Design challenges of the next
decades.

In the last years, self-assembled nanoscale
architectures~\cite{yan2003} emerged as a promising technology due
their tera/peta scale of integration, defect tolerance and huge
potential computing capabilities. These technologies are certainly
still at their early stage of development, however different
laboratory demos and proof-of-concepts have been
presented~\cite{patwardhan2004, patwardhan2006, patwardhan2006_1,
pistol2009}. For a complete survey on self-assembled architectures and
how they compare against classical many-core systems in terms of
 technologies, fault tolerance, performance and software tools see
 also~\cite{Future}.
%, patwardhan2006
The main idea behind this approach is exploiting the physical
regularity and stability of DNA structures in order to create a
scaffold onto which nano-devices (e.g. nanowires and
CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be attached. This
can be achieved by designing appropriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold (see Figure~\ref{fig:nana}-a).  Of course, a detailed
description of the chemical properties involved is far beyond the
scope of this paper (see also~\cite{braun1998, seeman1999}).

In this work we present \disr{} (Distributed Segment Routing), a distributed
approach focused on addressing the three main challenges that this new fabrication process introduces
in nanoscale Network-on-Chip Design: \emph{(i) limited node complexity}, \emph{(ii) large scale
randomness} and \emph{(iii) high defect rates}.  
The \emph{limited node complexity} aspect is directly
related to the use of complementary DNA tags in order to place circuit
components. The traditional CMOS process introduces complexity 
with larger photolithography masks: more complex (larger) circuits
will only require larger masks. Conversely, a self-assembly process achieves
complexity by increasing the number of unique DNA tags, since having more
different tags means having more control on component placement. Ideally, by specifying a single
and unique tag for each nano device terminal, we could exactly
choose where each component would be placed in the design, but the number of DNA
symbols forming the DNA sequence is limited (sequence of 4 nucleobases
G,A,T,C) and so creating many different tags (of a prefixed length) would
mean make them more similar to each other, increasing the probability
of incorrect/partial matching. To avoid this problem 
the number of unique tags must be limited, this limits also the
complexity available at each node.
In this work a budget of about $10,000$ CNFETs per node
has been assumed, as estimated in~\cite{liu_jetcs}.
\emph{Large scale randomness} is the other fundamental condition of
self-assembled technology: DNA tags allow control placement inside the
node grid, but there is no control over the placement of these grids
in the whole network. As a consequence, other typical
properties of regular networks cannot be guaranteed, e.g. being
connected to a fixed number of neighbors, having a determined
orientation and so on.
Finally, \emph{defect rates}: while they are very hardly tolerated in mask
based top-down design, the same nature of a bottom-up self-assembly
 process cannot assume such a deterministic device placement
process, thus defect tolerance is more a design requirement
than an exception to be avoided.

@@

\section{Background and Contribution}
\label{sec:related_works}
Several works addressed the problem of topology agnostic deadlock
freedom introducing virtual channels~\cite{sancho2002, skeie2002,
skeie2004, koibuchi2003}, but while they show interesting
performances, the low resources available to prevent deadlocks makes them not suitable for
the DNA self-assembled networks we are assuming as
scenario.
Other topology independent algorithms~\cite{schroeder1991,
koibuchi2001, cherkasova1996} do not require any virtual channel but
exploit the concept of ``turn prohibition''.
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploit the creation of a spanning tree of the
topology, placing bidirectional restrictions to avoid that a packet
can traverse a given link in both up and down directions.
The hierarchical nature of this approach can lead to an uneven traffic
distribution, with many packets traversing upper links (near to the
root), but this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010,  park2006, ebrahimi2013}.

>>>>

Exploring long term alternatives to the CMOS technology is gaining
more and more relevance as the scaling trend of such devices keeps
introducing new challenges: power density, defect tolerance, testing
costs and wire delays are only a few of the many critical aspects
involved~\cite{itrs13}. 
While software parallelism and multicore
approaches~\cite{Kavi2012243,Sibai2011958} are partially mitigating the impact of such
%approaches~\cite{horowitz2004,powell2009} are partially mitigating the impact of such
constraints on performances, it is likely that a growing computing demand will
eventually need even more radical architectural modifications and 
paradigms to address the Computer Design challenges of the next
decades.

In the last years, self-assembled nanoscale
architectures~\cite{yan2003} emerged as a promising technology due
their tera/peta scale of integration, defect tolerance and huge
potential computing capabilities. These technologies are certainly
still at their early stage of development, however different
laboratory demos and proof-of-concepts have been
presented~\cite{patwardhan2006_1, pistol2009}. For a complete survey on self-assembled architectures and
%presented~\cite{patwardhan2004, patwardhan2006, patwardhan2006_1, pistol2009}. For a complete survey on self-assembled architectures and
how they compare against classical many-core systems in terms of
 technologies, fault tolerance, performance and software tools see
 also~\cite{Future}.
%, patwardhan2006
The main idea behind this approach is exploiting the physical
regularity and stability of DNA structures in order to create a
scaffold onto which nano-devices (e.g. nanowires and
CNFETs~\cite{bachtold2001, cui2001}) can be attached. This
%CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be attached. This
can be achieved by designing appropriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold (see Figure~\ref{fig:nana}-a).  Of course, a detailed
description of the chemical properties involved is far beyond the
scope of this paper (see also~\cite{seeman1999}).

In this work, we present \disr{} (Distributed Segment Routing), a distributed
approach focused on addressing the three main challenges that this new fabrication process introduces
in nanoscale Network-on-Chip Design: \emph{(i) limited node complexity}, \emph{(ii) large scale
randomness} and \emph{(iii) high defect rates}.  
The \emph{limited node complexity} aspect is directly
related to the use of complementary DNA tags in order to place circuit
components. The traditional CMOS process introduces complexity 
with larger photolithography masks: more complex (larger) circuits
will only require larger masks. Conversely, a self-assembly process achieves
complexity by increasing the number of unique DNA tags, since having more
different tags means having more control on component placement. Ideally, by specifying a single
and unique tag for each nano device terminal, we could exactly
choose where each component would be placed in the design, but the number of DNA
symbols forming the DNA sequence is limited (sequence of 4 nucleobases
G,A,T,C) and so creating many different tags (of a prefixed length) would
mean make them more similar to each other, increasing the probability
of incorrect/partial matching. To avoid this problem 
the number of unique tags must be limited, this limits also the
complexity available at each node.
In this work a budget of about $10,000$ CNFETs per node
has been assumed, as estimated in~\cite{liu_jetcs}.
\emph{Large scale randomness} is the other fundamental condition of
self-assembled technology: DNA tags allow control placement inside the
node grid, but there is no control over the placement of these grids
in the whole network. As a consequence, other typical
properties of regular networks cannot be guaranteed, e.g. being
connected to a fixed number of neighbors, having a determined
orientation and so on.
Finally, \emph{defect rates}: while they are very hardly tolerated in mask
based top-down design, the same nature of a bottom-up self-assembly
 process cannot assume such a deterministic device placement
process, thus defect tolerance is more a design requirement
than an exception to be avoided.

@@

\section{Background and Contribution}
\label{sec:related_works}
Several works addressed the problem of topology agnostic deadlock
freedom introducing virtual channels~\cite{Abbas2014, skeie2004, koibuchi2003}, but while they show interesting
%freedom introducing virtual channels~\cite{sancho2002, skeie2002, skeie2004, koibuchi2003}, but while they show interesting
performances, the low resources available to prevent deadlocks makes them not suitable for
the DNA self-assembled networks we are assuming as
scenario.
Other topology independent algorithms do not require any virtual channel but
exploit the concept of ``turn prohibition''.
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploit the creation of a spanning tree of the
topology, placing bidirectional restrictions to avoid that a packet
can traverse a given link in both up and down directions.
The hierarchical nature of this approach can lead to an uneven traffic
distribution, with many packets traversing upper links (near to the
root), but this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
%topologies by limiting the number~\cite{duato1997, gomez2004, koibuchi2008} or the
topologies by limiting the number~\cite{ gomez2004, koibuchi2008} or the
location~\cite{flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
%dynamically recover defects as in~\cite{constantinides2006, kohler2010,  park2006, ebrahimi2013}. 
dynamically recover defects as in~\cite{ebrahimi2013}.
